---
title: "How to Extract Fromate from Excel"
author: "Dr.lumine"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: inline
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-"
)
```

## Acknowledgement

The tutorial thanks to: <https://www.r-bloggers.com/2018/05/tidying-messy-excel-data-tidyxl/>

Introduce: 

  - `library(tidyxl)` 
  - `library(upivotr)`

##installation

```{r}
setwd("~/Dropbox/Projects/Learning - R general/01 - Read Excel")
#devtools::install_github('nacnudus/unpivotr')
library(tidyxl)
library(unpivotr)

library(tidyverse)
```

```{r echo=TRUE}
cx <- list.files("data", "Excel", full.names = T) %>%
  xlsx_cells()
cx.colors <- cx %>%
  dplyr::filter(sheet == "a.colors")

names(cx.colors)
print("sheet names:")
list.files("data", ".xlsx", full.names = T) %>% xlsx_sheet_names()
```


## Formating
```{r}
cx.colors$local_format_id
```
So two things dictating the formate of a cx object `syle_format` and `local_format_id`, the later is an id that stores the style imformation. Later will show you how this is extracted in a completely different file

> To look up the local formatting of a given cell, take the cell's `local_format_id` value `(my_cells$Sheet1[1, "local_format_id"])`, and use it as an index into the format structure. E.g. to look up the font size, `my_formats$local$font$size[local_format_id]`. To see all available formats, type str(my_formats$local).

### Background Color
Strangely background color is stored in `fgColor$rgb` not anywhere else
```{r}
fx <- list.files("data", ".xlsx", full.names = T) %>%
  xlsx_formats()
```
The best way to do this is to through a function
```{r}
id_to_format <- function(x){
  fx$local$fill$patternFill$fgColor$rgb[x]
}
cx.colors %>%
  behead("up", "headler", character) %>%
  mutate(color = id_to_format(local_format_id)) %>%
  pivot_wider(id_cols = c(row),names_from = "headler", values_from = "color")
```

### deal with tables floating out of nowhere
```{r}
cx.b <- filter(cx, sheet == "b.floating tables")
cx.b %>% 
  rectify()
cx.b %>%
  behead("up", "header", character) %>%
  distinct(header)
```
### Merged
```{r}
cx.c <- dplyr::filter(cx, sheet == "c.floating merged cells")
cx.c %>%
  rectify
cx.c %>%
  behead("left-up", "merged", character) %>%
  unpivotr::pack() %>% # pack second so you can creat headlers
  select(row, merged, value) %>%
  unpivotr::unpack(value) # unpack later so you can use that
```
- row 19-20 is gone 
- it is gone when you `behead` everything
- the title we created from "up-left" because it has always been an up left, made sure everything is beheaded.

So in short it would be impossible to umerge cells if there is no other value besides it.

> Do a merged cell always have value placed on top left? 

The answer is yes try place a value in excel, merge outside, unmerge it again, 
the location turn.

```{r}
cx.d <- cx %>% filter(sheet == "d.paralle merged cells")

## Value layer
cx.d %>%
  unpivotr::pack() %>% 
  rectify(value = value)
## Formating layer
cx.d %>%
  rectify(value = local_format_id)
fx <- list.files("data", ".xlsx", full.names = T) %>%
  xlsx_formats() 
fx
```
Just saying that even though the merged cells looks as if there were nothing but still underneath is formatting. 

It is not because it can recognise merged cells, but because your merged cells has borders..... 

Boarder is identified formate

# Update: use data tree to visualise which data
### Explore formate list 
```{r}
#install.packages("data.tree")
library("data.tree")
fx %>% map(names)
fx$local$border

fx$local$border$right$color$rgb
fx$local$border$top$color$rgb
fx$local$border$left$color$rgb
fx$local$border$bottom$color$rgb

```

```{r}
library(networkD3)
URL <- paste0(
        "https://cdn.rawgit.com/christophergandrud/networkD3/",
        "master/JSONdata//flare.json")

## Convert to list format
Flare <- jsonlite::fromJSON(URL, simplifyDataFrame = FALSE)

# Use subset of data for more readable diagram
Flare$children = Flare$children[1:3]

diagonalNetwork(List = Flare, fontSize = 10, opacity = 0.9)
```
## How to Visualise a Nested List? 
[stack Visualise object as tree](https://stackoverflow.com/questions/51608378/visualise-object-in-r-as-tree)

```{r}
x <- list(
  id = 1,
  status = "active",
  coord = list(phi=0, theta=1, r=1),
  mt = NULL
)
depth <- function(x) ifelse(is.list(x), 1 + max(sapply(x, depth)), 0)

toTree <- function(x) {
  d <- depth(x)
  if(d > 1) {
    lapply(x, toTree)
  } else {
    children = lapply(names(x), function(nm) list(name=nm))
  }
}
dt <- FromListSimple(toTree(x), nodeName = "x")
plot(dt)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
pluck(fx, 1) %>% names()
depth(fx)
vec_depth(fx)
pluck(fx, 2)
length(fx)

fx %>% unlist() %>% length()

get_list <- function(x){
        while(is.list(x)){
          for(i in 1:length(x)){
            xi <- pluck(x, i)
          }
        }
        xi
}
fx %>% unlist() %>% names() %>% str_remove_all("[0-9]+") %>%
  unique() %>%
  data.frame() %>%
  filter(grepl("area",`.`))
```

```{r}
test_reg <- function(re){
  str_view_all(
    "style.border.diagonal.color.theme.Normal
    local.font.strike
    local.font.color.indexed", re)
}
test_reg("(?<=\\.)\\w+")
test_reg("\\w+\\.\\w+")
test_reg("(?<=\\.)\\w+\\.\\w+(?=\\.)")
```
# Give in to `openxl`
```{r}
wb<- list.files("data", ".xlsx", full.names = T) %>%
  loadWorkbook()

getStyles(wb) %>% class()
wbStyle <- getStyles(wb)
wbStyle %>% map(names)
wbStyle %>% length
wbStyle %>% depth()
```

```{r}
# NOT RUN {
## Create a new workbook
wb <- createWorkbook()

## Add a worksheet
addWorksheet(wb, "Sheet 1")
addWorksheet(wb, "Sheet 2")

## Merge cells: Row 2 column C to F (3:6)
mergeCells(wb, "Sheet 1", cols = 2, rows = 3:6)

## Merge cells:Rows 10 to 20 columns A to J (1:10)
mergeCells(wb, 1, cols = 1:10, rows = 10:20)

## Intersecting merges
mergeCells(wb, 2, cols = 1:10, rows = 1)
mergeCells(wb, 2, cols = 5:10, rows = 2)
mergeCells(wb, 2, cols = c(1, 10), rows = 12) ## equivalent to 1:10 as only min/max are used
# mergeCells(wb, 2, cols = 1, rows = c(1,10)) # Throws error because intersects existing merge

## remove merged cells
removeCellMerge(wb, 2, cols = 1, rows = 1) # removes any intersecting merges
mergeCells(wb, 2, cols = 1, rows = 1:10) # Now this works

## Save workbook
# }
# NOT RUN {
saveWorkbook(wb, "data/mergeCellsExample.xlsx", overwrite = TRUE)
# }
```

# Ideas for production
Okayy...this seems fun... just I can not possible find a "perfect" solution. 
I guess there were two ways to experiment with: 
  a. use the neighboring values to determine if a 
  b. format 
    b.a general format
    b.b boarder forma - which is most likely to be unreliable. 
  
A little CBT extraction idea: 
find where splitter is, then find row, in the same row, find maximum column that are not NULL. Then you locate where our CBT is!!

To find CBT port value... just look around neighbor and `length()` not Blank. or literally just `col - 1`


Extracting FDPs links should be much easier??? As done `behead(. "right-up", "outgoing_fdp)` OMG this is OP!! 
We just need a bit analytic around where to cut, it is li~terally just `filter()`

My data looks like this:
```
___________________
ENCLOSURE fdp
-------------------
Subs       |   |  |
xxxx       |   |  |
-------------------
(a lot  
    of 
blank space)
--------
fdp|   |------------X----------------------------------------> | X |~ cbt
   |   |------------X------------------------> | X |~ cbt      |   |
   |   |------c------> |fdp|   | #             |   |           |   |~ cbt 
   |   |------c------> |   |   | #             |   |.          |   |
   |   |.                                      |   |.
   |   |.
   |   |.
   |   |.
   |   |.
--------
```
If we consider spine node and edge all together is totally fun ~ we visualize things as it is... it is just going to have much-root. 

## Do merged cells have special formate? 
`t()` and `flatten()` really helps a lot to identify which part of the code is different. 
```{r}
fx %>% flatten() %>% data.frame -> ftb

  cx.c %>% rectify(value = local_format_id)
cx.d %>% rectify(value = local_format_id)


c(12, 13, 14) %in% 1:nrow(ftb)

ftb[1:nrow(ftb) %in% c(12, 13, 14, 25), ] %>% 
  t() %>%
  data.frame()
```
So merged cells will share some same formate such as font, background color (exceptions are boarders)... however cells share teh same formate does not means they are mereged... make this unreliable way of identifying a mreged cell.

